{\fontfamily{ptm}\selectfont Reading Notes}

### 2025-05-27 Notebook LaTEX/Formalized Notes/Notes 2025 03/main.tex
>w<, changes

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    \par
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \todo{[?]} and Nesterov \todo{[?]} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \todo{[?]} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique pioneer early by ??? \todo{[?]}.
    Necoara et al. \cite{necoara_linear_2019} showed that there exists an optimal restarting interval to achieve fast line convergence rate for all functions with quadratic growth condition. 
    \par
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice.

@@

@@

        The following definition can capture monotone variants of FISTA with line search, including backtracking strategies. 

@@

@@

            A simpler upper bound is more useful in practice. 

@@

            To see (i), use Lemma \ref{lemma:gmapg-seq-bnd} and its remark to bound $(\beta_k)_{k \ge 1}$ and then, apply Theorem \ref{thm:gmapg-ls-convergence} because the assumptions of $x^+, (\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ is the same. 

@@

@@

@@

@@

@@

                            $x_k, \eta_{k + 1}, G_k := \mathbf M(f, g, \tilde x_k, x_{k - 1}, \eta_k, G_k)$.
                        \IF{$G_k < \epsilon \textbf{ or } (\mathbf E_\chi  \textbf{ and } k > N_{\min})$}

@@

                \caption{GMAPG with Chambolle's backtracking}
                \end{algorithmic}\label{alg:gmapg}

@@

    % SECTION | EXAMPLES OF GMAPG IN THE LITERATURE

@@

    \section{Restarting with function values}

>>>>

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    \par
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \cite{beck_fast_2009-1} and Nesterov \cite[2.2.32]{nesterov_lectures_2018} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \cite{calatroni_backtracking_2019} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique discussed in Necoara et al. \cite{necoara_linear_2019}.
    They showed that there exists an optimal restarting interval to achieve fast line convergence rate for all functions with quadratic growth condition. 
    The method of restarting was improved to be parameter free while still retaining a fast linear convergence rate, see Alamo et al. \cite{alamo_restart_2019} and Aujol et al. \cite{aujol_parameter-free_2024}. 
    \par
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice.

@@

        Firstly, recall the definition of Bregman divergence $D_f(x, y)$ from Definition \ref{def:bregman-div} for a differentiable function $f:\RR^n \rightarrow \overline \RR$. 

@@

        The following definition captures several monotone variants of FISTA with line search,  or backtracking strategies.
        In this section, we give convergence results for a unified formulation of monotone accelerated proximal gradient methods with line search or backtracking enhancements. 
        Definition \ref{def:generic-mapg-ls} gives a unified view of several combined heuristics. 
        Theorem \ref{thm:gmapg-ls-convergence} gives a generic description for the convergence rate. 
        Lemma \ref{lemma:gmapg-seq-bnd} specializes the sequence, attains the lowest upper bound. 
        Theorem \ref{thm:gmapg-specialized-convergence} gives a concrete $\mathcal O(1/k^2)$ upper bound for the optimality gap and normed gradient mapping on the last iteration. 

@@

            The above theorem is improved from Alamo et al. \cite{alamo_restart_2019}. 

@@

            The techique of the proof we used here is very similar to Güler \cite[Lemma 2.2]{guler_new_1992}.
            A simpler upper bound is more practical. 

@@

            To see (i), use Lemma \ref{lemma:gmapg-seq-bnd} and its remark to bound $(\beta_k)_{k \ge 1}$ and then, apply Theorem \ref{thm:gmapg-ls-convergence} because the assumptions of $x^+, (\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ suit. 

@@

            Algorithm \ref{alg:armijo-ls} performs a step of Armijo line search and a step of accelerated proximal gradient. 
            The function can be used for each iteration in the inner loop of the algorithm. 
            Here are the explanations for all its input parameters: 
                \item $f, g$ are functions satisfying Assumption \ref{ass:standard-fista}. 
                \item $x, v$ are the $x_k, v_k$ iterates in Definition \ref{def:generic-mapg-ls}. 
                \item $\alpha$ are the current $\alpha_k$ in Definition \ref{def:generic-mapg-ls}. 
                \item $L$ is the estimate of the Lipschitz constant of $f$ passed in by the inner loop. 
            Iterates $x^+, y^+$ and parameters $\alpha^+, L^+$ are returned to the callers at the end. 

@@

            Algorithm \ref{alg:chambolle-btls} attempts to decrease the Lipschitz estimate $L_k$ for $f$ in an iteration of the inner loop. 
            The above implementations were adapted and simplified from Chambolle et al. \cite{calatroni_backtracking_2019}. 
            It takes in additional parameters $L_{\min}, \rho$ compared to Algorithm \ref{alg:armijo-ls}. 
            Here are their explanations: 
                \item $L_{\min}$ determines a lower bound of Lipschitz estimates. It's the lowest value of an estimate $L_k$ allowed. It increases stability of the algorithm by preventing unnecessary triggering a line search routine to recovers from an underestimated $L_k$ that doesn't satisfy the Lipschitz smoothness condition for $f$ at the current iterate. 
                \item $\rho \in (0, 1)$ is the decay ratio. It's use to shrink the current estimate of $L$ and produce $L^+$ at the start of the forloop before verifying the smoothness condition. 

@@

            Algorithm \ref{alg:beck-mono} is a subroutine for asserting monotone condition on function value. 
            The parameter $G$ has no actual usage besides making it compatible with Algorithm \ref{alg:nes-mono} in the context of Algorithm \ref{alg:gmapg}. 
            The input $\tilde x$ is the candidate iterate produced by FISTA without monotone constraints and $x$ is the previous iterates $x_{k - 1}$ in the inner loop. 

@@

            The above Algorithm \ref{alg:nes-mono} implements and adapts Nesterov's monotone scheme from Nesterov \cite[2.2.32]{nesterov_lectures_2018} for GMAPG. 
            In addition to Algorithm \ref{alg:beck-mono}, $\eta$ is a new input parameter and $G$ has a significance role. 
            $\eta$ is a stepsize parameter for weakly convex objective $F = f + g$ satisfying Assumption \ref{ass:sum-of-wcnvx}. 
            $G$ is the norm of the gradient mapping updated at at $\hat y$ which will be returned to the inner loop to verify exit conditions. 

@@

                            $x_k, \eta_{k + 1}, G_k^+ := \mathbf M(f, g, \tilde x_k, x_{k - 1}, \eta_k, G_k)$.
                        \IF{$G_k^+ < \epsilon \textbf{ or } (\mathbf E_\chi  \textbf{ and } k > N_{\min})$}

@@

                \caption{GMAPG with Chambolle's backtracking}\label{alg:gmapg}

@@

            The above Algorithm \ref{alg:gmapg} is an implementation of GMAPG in Definition \ref{alg:gmapg}. 
            Parameters $r, \rho$ are chosen in the discretion of the practitioners. 
            For example, we chose $r= 0.4, \rho = 2^{1/1024}$.

@@

    \section{Restarting with function values for linear convergence}
        In this section, we show that restarting GMAPG with a well suited conditions yields fast linear global convergence. 
        We adapt and improve prior theories from Alamo \cite{alamo_restart_2019} for our GMAPG method. 
        Let's introduce our first set of restart conditions which denote it by $\mathbf E_{\chi}^{a}$. 
        $\mathbf E_{\chi}^{a}$ only uses function values on iterates available to the inner loop to determine the exit conditions, and it achieves global fast linear convergence. 
        The inner loop algorithm is described in Algorithm \ref{alg:gmapg}, let $k$ denote the inner loop counter and define $m = \lfloor k/2 \rfloor + 1$ the conditions are: 
            \mathbf E_{\chi}^a \iff 
            f(x_m) - f(x_k) \le \exp(-1)(f(x_0) - f(x_m)). 
            \STATE{\textbf{Input: }\begin{tabular}{ll}
                $M \in \N$ & Integer\\
                $L:=1$ & $L > 0$\\ 
                $r:=0.5$ & $r \in (0, 1)$ \\ 
                $\rho := 2^{1/1024}$ & $\rho \in (0, 1)$ 
            \end{tabular}}
            \STATE{$n_0 := 0$.}
            \STATE{$z_1, p_0 := \textbf{GMAPG}(f, g, x_{-1}, r_j, \rho, N_{\min}=n_0, N=M, \epsilon, \textbf{L}, \textbf{M}, \textbf{E}_{\chi}^a).$}
            \STATE{$n_1:=p_0. $}
            \STATE{$M := M - n_1$}
            \FOR{$j = 1,2, \ldots, M$}
                \STATE{$z_{j + 1}, p_j := \textbf{GMAPG}(f, g, z_j, r, \rho, N_{\min}=n_j, N=M, \epsilon, \textbf{L}, \textbf{M}, \textbf{E}_{\chi}^a).$}
                \STATE{$M := M - p_j$}
                \IF{$f(z_{j}) - f(z_{j + 1}) > \exp(-1)(f(z_{j - 1}) - f(z_{j}))$}
                    \STATE{$n_{j + 1} := 2p_j$.}
                \ELSE
                \STATE{$n_{j + 1} := p_j$.}
            \end{algorithmic}\caption{Linear convergence restarted GMAPG}\label{alg:linear-rgmapg}
        Algorithm \ref{alg:linear-rgmapg} implements a restarted GMAPG with condition $\textbf{E}_{\chi}^a$ and, it has a fast linear convergence rate. 
        The following observation on it is crucial to take notice for the convergence rate. 
        $n_j$ is used as the lower bound $N_{\min}$ for the GMAPG inner loop described in Algorithm \ref{alg:gmapg} at iteration $j$ and, $p_j$ is the actual iteration underwent by the inner loop and, it has $p_j \ge n_{j}$. 
        $n_{j + 1}$ is the minimum iteration for the next execution of GMAGP, and it has $n_{j + 1} = p_{j}$  if the ``if statement" isn't triggered otherwise $n_{j + 1} = 2p_{j}$. 
        Consequently, it has either $p_{j + 1} \ge n_{j + 1} = 2p_j$ or $p_{j + 1} \ge n_{j + 1} = p_j$. 
        We now introduce the following notations for the convergence proof. 
        Since the restarted GMAPG consist of executions of an outer loop in Algorithm \ref{alg:linear-rgmapg} in variable $j$ and, an inner loop in Algorithm \ref{alg:gmapg} in variable $k$, we denote $z_{j, k}$ for the iterates $x_k$ in the inner loop and $z_j$ in the outer loop together. 
        We make the rule that $z_{j + 1, -1} = z_{j, p_j}$ for compatibility of the theorems in the previous sections. 
        For example in line 7 at the $j$ iteration, the inner loop returns $x_{p_j}$ as the last iterate of the inner loop and, it's assigned to $z_{j + 1}$ by the outer loop. 
        In that case we denote $z_{j + 1, -1}$ as the initial iterate pass into the inner loop for the $j + 1$ iteration, and $z_{j, p_j}$ would denote the last iteration before exiting the inner loop for the $j$ th iteration of the outer loop. 
        Therefore, $z_{j + 1, -1} = z_{j, p_j}$ and all iterates generated by the inner loop is the finite sequence $(z_{j, k})_{k = 0}^{p_j}$. 

        \begin{lemma}[bounds on inner iteration counts]
            ...
        \end{lemma}
        \begin{proof}

        \end{proof}
        \begin{lemma}[bounds on outer iteration counts]
            ...
        \end{lemma}

        \begin{theorem}[bounds on the total iterations]
            ...
        \end{theorem}
